Long-Term Reproducibility through Replication:\\ the ReScience Initiative

### 2017-06-07 draft.tex
Extend dicussion, remove reproducibility is harder than you think section

\section*{Replication and reproduction in computer-aided research}

With the increasing recognition of the replication crisis as a problem
for scientific research, it has become a major subject of debate, but
unfortunately no common terminology has emerged so far. One reason for
this diverse use of terms is that each field of research has its own
specific technical and social obstacles on the road to published
results and findings that can be verified by other scientists. Here we
briefly summarize the obstacles that arise from the use of computers
and software in scientific research, and introduce the terminology we
will use in the rest of this article.

@@

Strictly speaking, reproducibility is defined in the context of
identical computational environments. However, useful scientific
software is expected to be robust with respect to certain changes in
this environment. A computer program that produces different results
when compiled using different compilers, or run on two different
computers, would be considered suspect by most practitioners, even if
it were demonstrably correct in one specific environment. Ultimately
it is not the software that is of interest for science, but the models
and methods that it implements. The software is merely a vehicle to
perform computations based on these models and methods. If results
depend on hard to control implementation details of the software,
their relation to the underlying models and methods becomes unclear.

@@

\section*{Reproducibility is harder than you think}

\citeauthor{Mesnard:2016} explains in \citep{Mesnard:2016} that performing
reproducible and replicable simulations in computational fluid dynamics is harder
than you think. In fact, there are several different reasons for
non-reproducibility that, for the most part, are specific to each
scientifc domain. However, from our publishing experience, there are also some
common behaviors that can be identified. Missing code and/or data, unknown
dependencies, inaccurate or imprecise description appears to be shared
characteristic of a lot of non-reproducible work.\\

To be extended...

@@

\begin{tcolorbox}[breakable, pad at break*=1mm,
                  colback=black!2.5, arc=0pt, outer arc=0pt, boxrule=0pt]
\begin{footnotesize}
\textbf{Box 1.} Authors having  published in Rescience explain their motivation.\\

@@

%% ReScience already published 4 articles and as shown above, the original
%% motivations of these authors are all different and this might become even more
%% obvious and diverse with future publications. But, beyond these motivations,
%% publishing in ReScience may be especially important for students since this
%% represent a unique opportunity to show the community a given student is able to
%% read a scientific article, to have a deep understanding of it, to write a new
%% implementation and to eventually write a scientific article describing his/her
%% work. Although the ReScience publishing model is a bit different from other
%% academic journals, it can give students a first experience at peer-reviewed
%% scholarly publishing, including meeting standards of scientific rigor and
%% addressing reviewers' comments.  Furthermore, publishing in ReScience is a way
%% to actively contribute to open science while adding to one's publication
%% record.

@@

\begin{figure}
  \includegraphics[width=1.0\columnwidth]{CoScience}
  \caption{\textbf{A} The ReScience publication chain starts from an
    original research by authors A, published in a journal, in
    conference proceedings, or as a preprint. This article constitutes
    the base material for authors B who attempt to replicate the work
    based on its description. Success or failure to replicate is not a
    criterion for acceptance or rejection, even though failure to
    replicate requires more precaution to ensure this is not a
    misunderstanding or a bug in the new code. After review, the
    replication is published, and feedback is given to original
    authors (and editors) to inform them the work has been replicated
    (or not). \textbf{B} The CoScience proposal would require the
    replication to happen \textit{before} the actual publication. In
    case of failure, nothing will be published. In case of success,
    the publication will be endorsed by authors A and authors B with
    identified roles and will be certified as reproducible because it
    has been replicated by an independent group.}
\end{figure}

@@

% -> Short-term vs long-term reproducibility
The code accompanying a ReScience replication ensures only short-term
reproducibility. Even if a piece of code is written following best practices,
and then reviewed and tested, it suffers from the same problem as any other piece of
code, i.e. it depends on a software stack whose stability is not guaranteed in
the long term. \citep{Mesnard:2016} illustrated this problem very clearly when
they tried to reproduce their own work performed two years earlier. Even though Barba's
group is committed to reproducible research practices, they did not escape the many
problems one can face when trying to re-run a piece of code. This means the
newly produced code for ReScience will likely be obsolete at some point in the
future. It can be a matter of months or years, or even decades, but ultimately, obsolescence is inevitable. Therefore, the long-term value of a ReScience
publication is not the actual code but the accompanying article. If you
consider the original article with missing or incomplete information and the
new article with missing and checked information, you have a complete and
consistent description of the original work. This means than 5, 10 or 20 years
from now, one should be able to replicate the work thanks to these two
articles. Of course, the new code can also help, but the true value of a
replication is the accompanying article.

>>>>

\section*{Replication and reproduction}

With the increasing recognition of the replication crisis as a problem
for scientific research, it has become a major subject of debate, but
unfortunately no common terminology has emerged so far. One reason for
this diverse use of terms is that each field of research has its own
specific technical and social obstacles on the road to published
results and findings that can be verified by other scientists. Here we
briefly summarize the obstacles that arise from the use of computers
and software in scientific research, and introduce the terminology we
will use in the rest of this article.\\

@@

Strictly speaking, reproducibility is defined in the context of
identical computational environments. However, useful scientific
software is expected to be robust with respect to certain changes in
this environment. A computer program that produces different results
when compiled using different compilers, or run on two different
computers, would be considered suspect by most practitioners, even if
it were demonstrably correct in one specific environment. Ultimately
it is not the software that is of interest for science, but the models
and methods that it implements. The software is merely a vehicle to
perform computations based on these models and methods. If results
depend on hard to control implementation details of the software,
their relation to the underlying models and methods becomes unclear.\\

@@

@@

\begin{tcolorbox}[breakable, pad at break*=1mm,
                  colback=black!2.5, arc=0pt, outer arc=0pt, boxrule=.25pt]
\begin{footnotesize}
\textbf{Box 1.} Authors having  published in Rescience explain their motivation.\\

@@

%% ReScience already published 4 articles and as shown above, the original
%% motivations of these authors are all different and this might become even more
%% obvious and diverse with future publications. But, beyond these motivations,
%% publishing in ReScience may be especially important for students since this
%% represent a unique opportunity to show the community a given student is able to
%% read a scientific article, to have a deep understanding of it, to write a new
%% implementation and to eventually write a scientific article describing his/her
%% work. Although the ReScience publishing model is a bit different from other
%% academic journals, it can give students a first experience at peer-reviewed
%% scholarly publishing, including meeting standards of scientific rigor and
%% addressing reviewers' comments.  Furthermore, publishing in ReScience is a way
%% to actively contribute to open science while adding to one's publication
%% record.
%% \section*{Reproducibility is harder than you think}

%% \citeauthor{Mesnard:2016} explains in \citep{Mesnard:2016} that performing
%% reproducible and replicable simulations in computational fluid dynamics is harder
%% than you think. In fact, there are several different reasons for
%% non-reproducibility that, for the most part, are specific to each
%% scientifc domain. However, from our publishing experience, there are also some
%% common behaviors that can be identified. Missing code and/or data, unknown
%% dependencies, inaccurate or imprecise description appears to be shared
%% characteristic of a lot of non-reproducible work.\\

%% To be extended...

@@

\begin{figure}
  \includegraphics[width=1.0\columnwidth]{CoScience}
  \caption{\textbf{A} The ReScience publication chain starts from an
    original research by authors A, published in a journal, in
    conference proceedings, or as a preprint. This article constitutes
    the base material for authors B who attempt to replicate the work
    based on its description. Success or failure to replicate is not a
    criterion for acceptance or rejection, even though failure to
    replicate requires more precaution to ensure this is not a
    misunderstanding or a bug in the new code. After review, the
    replication is published, and feedback is given to original
    authors (and editors) to inform them the work has been replicated
    (or not). \textbf{B} The CoScience proposal would require the
    replication to happen \textit{before} the actual publication. In
    case of failure, nothing will be published. In case of success,
    the publication will be endorsed by authors A and authors B with
    identified roles and will be certified as reproducible because it
    has been replicated by an independent group.}
  \label{fig:coscience}
\end{figure}

@@

% -> Short-term vs long-term reproducibility

There are several different reasons for non-reproducibility. For the most part,
those reasons are specific to each scientifc domain but, from our publishing
experience, there are also some common behaviors that can be identified.
Missing code and/or data, unknown dependencies, inaccurate or imprecise
description appears to be shared characteristic of a lot of non-reproducible
work. If the code accompanying a ReScience replication can fix a lot of these
issue, it can nonetheless only ensure short-term reproducibility. Even if a
piece of code is written following best practices, and then reviewed and
tested, it suffers from the same problem as any other piece of code, i.e. it
depends on a software stack whose stability is not guaranteed in the long
term. \citep{Mesnard:2016} illustrated this problem very clearly when they
tried to reproduce their own work performed two years earlier. Even though
Barba's group is committed to reproducible research practices, they did not
escape the many problems one can face when trying to re-run a piece of code.
This means the newly produced code for ReScience will likely be obsolete at
some point in the future. It can be a matter of months or years, or even
decades, but ultimately, obsolescence is inevitable. Therefore, the long-term
value of a ReScience publication is not the actual code but the accompanying
article. If you consider the original article with missing or incomplete
information and the new article with missing and checked information, you have
a complete and consistent description of the original work. 5, 10 or 20 years
from now, anyone should be able to replicate the work thanks to these two
articles. Of course, the new code can also help, but the true value of a
replication is really the accompanying article.\\

One immediate and legitimate question is to wonder to what extent such
replication could be performed prior to the publication of the original
article. This would strongly reinforce a claim because a successful and
independent replication would have been performed prior to the publication. As
illustrated on figure \ref{fig:coscience}, this would require for a group A to
contact a group B and to send them their a draft of their original work (the
one that would be normally submitted to a journal) such that group B could run
a replication and confirm or infirm the work. In case of confirmation, a
certified article could be later published with both group as authors (each
group being identified according to their respective roles). However, if the
replication failed and the original work cannot be fixed, this would prevent
the publication. This model would improve the overall reproducibiloty and would
slow down considerably the pace of publication we're observing
today. Unfortunately, such scenario is highly improbable. The pressure to
publish is so strong and the incentive for doing replication so low that it
would most probably prevent such collaborative work. At this point, CoScience
is mostly an utopia for tomorrow.
